Different rules for binocular combination of luminance flicker in cortical and subcortical pathways

### 2023-01-20 manuscriptNature.Rmd
Created rmd file specific for Nature and added Nature referencing style.

---
title: "Different rules for binocular combination of luminance flicker in cortical and subcortical pathways"
author: "Federico G. Segala, Aurelio Bruno, Myat T. Aung, Alex R. Wade &  Daniel H. Baker"
date: "Department of Psychology, University of York, UK"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  pdf_document:
    toc: no
  word_document: default
  html_document: default
bibliography: references.bib
csl: elife.csl
---

@@

Thirty (20 females), twelve (7 females) and ten (3 females) adult participants, whose ages ranged from 18 to 45, were recruited for Experiments 1, 2 and 3 respectively. All participants had normal or corrected to normal binocular vision, and gave written informed consent. Our procedures were approved by the Ethics Committee of the Department of Psychology at the University of York (identification number 792).

The stimuli were two discs of achromatic flickering light with a diameter of 3.74 degrees, presented on a black background. The same stimuli were used for all three experiments. Four dark red lines were added around both discs to help with their fusion into one binocular disc (see insert in Figure \@ref(fig:pupildata)b for an example of the fused stimulus). The discs were viewed through a four-mirror stereoscope, which used front silvered mirrors to avoid internal reflections, and meant that participants saw a single fused disc. The use of a stereoscope allowed us to modulate the stimuli in three different ocular configurations: monocular, binocular, and dichoptic. Note that during monocular presentation of flicker, the unstimulated eye still saw the static (non-flickering) disc of mean luminance.

We implemented the model within a Bayesian framework using the Stan software [@Carpenter2017]. This allowed us to estimate group-level posterior parameter distributions for the weight of interocular suppression, $w$, and the other free model parameters $R_{max}$, $Z$ and $n$. The prior distributions for all parameters were Gaussian, with means and standard deviations of 1 and 0.5 for $w$ and $R_{max}$, and 5 and 2 for $Z$ and $n$, with these values chosen based on previous literature [@Meese2006; @Baker2012]. We sampled from a Student's t-distribution for the amplitudes in the pupillometry and EEG experiments, and from a Bernoulli distribution for the single trial matching data. The models were fit using the individual data across all participants, independently for each data set. Because we used coherent averaging across participants, the group average amplitudes are shifted vertically relative to the model predictions, which are based on hierarchical fits to the individual participant amplitudes (put another way, the model does not implement coherent averaging). However it is clear that the character of the model in all cases gives a good representation of the data, as can be seen in Figure \@ref(fig:modelfigure). We took posterior samples at over a million steps for each data set, using a computer cluster, and retained 10\% of samples for plotting.

## Preregistration, data and code availability

We initially preregistered our main hypotheses and analysis intentions for the first experiment. We then conducted a pilot study with N=12 participants, before making some minor changes to the stimulus (we added dim red lines to aid binocular fusion). We then ran the main experiment, followed by two additional experiments that were not preregistered. The preregistration document, raw data files, and experimental and analysis code, are available on the project repository: https://doi.org/10.17605/OSF.IO/TBEMA.

@@

The pupillometry results are summarised in Figure \@ref(fig:pupildata). The group average waveform for binocular presentation is shown in Figure \@ref(fig:pupildata)a. There is a substantial pupil constriction at stimulus onset, followed by visible oscillations at the flicker frequency (2Hz, see waveform at foot). The average Fourier spectrum is displayed in Figure \@ref(fig:pupildata)b, and shows a clear spike at 2 Hz, but no evidence of a second harmonic response at 4Hz. These results demonstrate that our paradigm can evoke measurable steady-state pupil responses at 2Hz.

@@

Figure \@ref(fig:pupildata)c shows contrast response functions in response to stimuli flickering only at 2Hz. Response amplitudes increased monotonically with target contrast, confirming that our paradigm is suitable for measuring contrast-dependent differences in response (to our knowledge this is the first time this has been demonstrated). The amplitude of the binocular condition (blue squares) is consistently greater than that of the monocular condition (red circles) across all target contrasts. A $2\times5$ repeated measures $ANOVA^2_{circ}$ [@Baker2021] comparing these conditions revealed a significant main effect of target contrast (F(`r monbinPstats$dfA`,`r monbinPstats$dfR`) = `r round(monbinPstats$FratioA,digits=2)`, $p$ < 0.001), a significant effect of condition (F(`r monbinPstats$dfB`,`r monbinPstats$dfR`) = `r round(monbinPstats$FratioB,digits=2)`, $p$ < 0.001), and a significant interaction (F(`r monbinPstats$dfAB`,`r monbinPstats$dfR`) = `r round(monbinPstats$FratioAB,digits=2)`, $p$ < 0.001). The dichoptic condition begins at a much higher amplitude, owing to binocular combination of the target and high (48%) contrast mask, and then increases slightly with increasing target contrast (main effect of target contrast: F(`r dichA2c$dfM`,`r dichA2c$dfR`) = `r round(dichA2c$Fratio,digits=2)`, $p$ < `r round(dichA2c$pval,digits=3)`).

In Figure \@ref(fig:pupildata)d, we plot responses to monocular target stimuli flickering at 2Hz, when the other eye viewed stimuli flickering at 1.6Hz (the red monocular-only data are replotted from Figure \@ref(fig:pupildata)c for comparison). When the 1.6Hz component had the same contrast as the target (the binocular cross condition, shown in purple) responses were facilitated slightly at low contrasts, and suppressed at the highest target contrasts (interaction between contrast and condition: F(`r monbinPstatsX$dfAB`,`r monbinPstatsX$dfR`) = `r round(monbinPstatsX$FratioAB,digits=2)`, $p$ < 0.001). When the 1.6Hz component had a fixed contrast of 48% (the dichoptic cross condition, shown in yellow), responses were suppressed slightly across the contrast range (interaction between contrast and condition: F(`r mondichPstatsX$dfAB`,`r mondichPstatsX$dfR`) = `r round(mondichPstatsX$FratioAB,digits=2)`, $p$ < 0.001).

Figure \@ref(fig:pupildata)e shows responses at 1.6Hz, for the same conditions, as well as for a condition in which a monocular stimulus flickered at 1.6Hz (grey circles). Surprisingly there again appears to be a slight facilitation effect in the binocular cross condition, particularly at lower contrasts. The dichoptic cross condition does not show clear modulation with target contrast.

Figure \@ref(fig:EEGdata) shows equivalent results, measured contemporaneously using EEG. Figure \@ref(fig:EEGdata)a shows the group average waveform for binocular presentation, and Figure \@ref(fig:EEGdata)b shows the Fourier spectrum for binocular presentation, both averaged across four posterior electrodes (_Oz_, _POz_, _O1_ and _O2_, marked on the inset scalp plots). Unlike for the pupillometry data, there are clear responses at both the first harmonic frequency (2Hz), and also the second harmonic frequency (4Hz). We therefore calculated contrast response functions at both first and second harmonic frequencies.

@@

When stimuli in both eyes flicker at 2Hz, the binocular responses at the first (Figure \@ref(fig:EEGdata)c) and second (Figure \@ref(fig:EEGdata)f) harmonics are substantially greater than the monocular responses, particularly at high contrasts. Analysis of variance on the complex values ($ANOVA^2_{circ}$) revealed a main effect of contrast (F(`r monbinEstats$dfA`,`r monbinEstats$dfR`) = `r round(monbinEstats$FratioA,digits=2)`, $p$ < 0.001) and an interaction effect (F(`r monbinEstats$dfAB`,`r monbinEstats$dfR`) = `r round(monbinEstats$FratioAB,digits=2)`, $p$ < 0.001), but no effect of condition ($p$ = `r round(monbinEstats$pvalB,digits=2)`) at the first harmonic, with a similar pattern of results obtained at the second harmonic. For the cross-frequency conditions (Figure \@ref(fig:EEGdata)d,g), there was no appreciable effect of adding a 1.6Hz component on the response at 2Hz or 4Hz (no effect of condition, and no interaction). Similarly, there were no clear interocular interactions between frequencies in the responses at 1.6Hz (Figure \@ref(fig:EEGdata)e) and 3.2Hz (Figure \@ref(fig:EEGdata)h). This pattern of results suggests that processing of temporal luminance modulations happens in a more linear way in visual cortex (indexed by EEG), compared with subcortical pathways (indexed by pupillometry), and shows no evidence of interocular suppression.

@@

Finally, we calculated the ratio of binocular to monocular responses across the three data types from Experiment 1. Figure \@ref(fig:BSratios) shows that these ratios are approximately $\sqrt2$ across the low-to-intermediate contrast range for all three data types. At higher contrasts, we see ratios of 2 or higher for the EEG data, but much weaker ratios near 1 for the pupillometry data. Note that the ratios here are calculated on a per-participant basis and then averaged, rather than being the ratios of the average values shown in Figures \@ref(fig:pupildata) and \@ref(fig:EEGdata). A $3 \times 5$ repeated measures ANOVA on the logarithmic (dB) ratios found a main effect of contrast (F(`r round(ratioaov$ANOVA$DFn[2]*gg[2],digits=2)`,`r round(ratioaov$ANOVA$DFd[2]*gg[2],digits=2)`) = `r round(ratioaov$ANOVA$F[2],digits=2)`, \emph{p} < `r round(ratioaov$ANOVA$p[2],digits=3)`), no effect of data modality (F(`r ratioaov$ANOVA$DFn[1]`,`r ratioaov$ANOVA$DFd[1]`) = `r round(ratioaov$ANOVA$F[1],digits=2)`, \emph{p} = `r round(ratioaov$ANOVA$p[1],digits=2)`), but a highly significant interaction (F(`r round(ratioaov$ANOVA$DFn[3]*gg[3],digits=2)`,`r round(ratioaov$ANOVA$DFd[3]*gg[3],digits=2)`) = `r round(ratioaov$ANOVA$F[3],digits=2)`, \emph{p} < `r round(ratioaov$ANOVA$p[3],digits=4)`).

@@

Results from the temporal frequency experiment are shown in Figure \@ref(fig:TFdata). Figure \@ref(fig:TFdata)a shows the Fourier spectra for responses to binocular flicker at 5 different frequencies (2, 4, 8, 16, and 30 Hz). From 2 to 16 Hz, clear signals are observed at each fundamental frequency, and typically also their higher harmonics (integer multiples of the fundamental). However, at 30 Hz (upper row), the responses recorded were not demonstrably above the noise baseline. Figure \@ref(fig:TFdata)b compares the monocular and binocular responses at each stimulation frequency. Here we replicate the substantial summation effect across frequencies up to and including 16Hz (Fig. \@ref(fig:TFdata)c), demonstrating that strong binocular facilitation in the EEG data of Experiment 1 cannot be attributed to our use of 2Hz flicker.

@@

The results from the matching experiment are shown in Figure \@ref(fig:matchingdata). Each data point indicates the contrast levels required in each eye that were perceptually equivalent to the binocular 24\% (red circles) and 48\% (blue circles) matching contrasts. At both matching contrasts, we see a very substantial increase in the physical contrast required for a monocular target (data points along the x- and y-axes), compared to a binocular target (points along the diagonal of x=y). For example with a 48\% match, the monocular targets required contrasts close to 100\%, whereas binocular targets required a contrast of around 50\%. The data points between these extremes also fall close to the predictions of a linear summation model (diagonal dotted lines), and are inconsistent with a winner-takes-all (or MAX) model (dashed lines). Overall, these matching results are consistent with the approximately linear summation effects observed in the EEG data of Experiment 1 (Figure \@ref(fig:EEGdata)c,f).

@@

We fitted a computational model to the data from Experiments 1 & 3 using a hierarchical Bayesian approach. The model behaviour is displayed in Figure \@ref(fig:modelfigure)e-h, with empirical data replotted in Figure \@ref(fig:modelfigure)a-d for comparison. In general, the model captures the key characteristics of the empirical data. Note that there are some minor discrepancies, which are a consequence of the hierarchical nature of the modelling. In brief, the model is fitted to the amplitudes for each participant, and group-level parameter estimates are derived based on these fits (see Table \@ref(tab:paramtable)). This procedure discards the phase information, whereas the empirical averages are coherently averaged across participants (retaining phase information). This explains the amplitude differences between model and data, particularly at low target contrast levels, but is of little consequence for the pattern of relative responses across conditions, which is our main focus here.

We were particularly interested in comparing the weight of interocular suppression across data sets. We therefore plot the posterior distributions for this parameter for all four data sets (see Figure \@ref(fig:modelfigure)i). The key finding is that the pupillometry results (green distribution) display a much greater weight of interocular suppression compared with the other data sets (grey, purple and yellow distributions). There is no overlap between the pupillometry distribution and any of the other three. All four distributions are also meaningfully below a weight of 1 -- the value that previous work using grating stimuli would predict [@Meese2006; @Baker2017], and the peak location of our prior distribution (black curve). These results offer an explanation of the empirical data: the strong interocular suppression for the pupillometry data is consistent with the weak binocular facilitation, and measurable dichoptic masking observed using that method. The weaker suppression for the other experiments is consistent with the near-linear binocular facilitation effects, and absent dichoptic masking.

@@

Classic studies investigating the neurophysiological architecture of V1 reported that cells in cytochrome-oxidase 'blobs' [@Horton1981; @Livingstone1984] are biased towards low spatial frequencies [@Tootell1988; @Edwards1995], and relatively insensitive to stimulus orientation [@Horton1981; @Livingstone1984; though see @Economides2011]. As the blob regions are embedded within ocular dominance columns [@Horton1981], they are also largely monocular [@Livingstone1984; @Tychsen2004]. More recent work has reported psychophysical evidence for unoriented chromatic [@Gheiratmand2013] and achromatic [@Meese2011] mechanisms, that also appear to be monocular. Our use of luminance flicker might preferentially stimulate these mechanisms, perhaps explaining why our EEG data show little evidence of binocular interactions. Indeed, our EEG results could potentially be explained by a model involving entirely non-interacting monocular channels, with the binocular facilitation effects we find (e.g. Figures \@ref(fig:BSratios) & \@ref(fig:TFdata)) owing to additivity of the electrophysiological response across independent monocular cells, rather than requiring binocular neurons. However our matching data (Figure \@ref(fig:matchingdata)), as well as everyday perceptual experience, indicates that luminance signals must be combined physiologically at some stage, and that this process still involves weak or absent interocular suppression.

@@

# Acknowledgements

Supported by BBSRC grant BB/V007580/1 awarded to DHB and ARW, and Wellcome Trust grant 213616/Z/18/Z to AB.

# Author contributions

**Federico Segala**: Methodology, software, formal analysis, investigation, data curation, writing - original draft, writing - review & editing, visualization. **Aurelio Bruno**: Conceptualization, writing - review & editing, supervision, project administration, funding acquisition. **Myat Aung**: Software, resources, writing - review & editing. **Alex Wade**: Conceptualization, methodology, resources, writing - review & editing, supervision, project administration, funding acquisition. **Daniel Baker**: Conceptualization, methodology, software, formal analysis, investigation, resources, data curation, writing - original draft, writing - review & editing, visualization, supervision, project administration, funding acquisition.

# Declaration of interests

>>>>

---
title: "Different rules for binocular combination of luminance flicker in cortical and subcortical pathways"
author: "Federico G. Segala, Aurelio Bruno, Myat T. Aung, Alex R. Wade &  Daniel H. Baker"
date: "Department of Psychology, University of York, UK"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  pdf_document:
    toc: no
  word_document: default
  html_document: default
bibliography: references.bib
csl: nature-neuroscience.csl
---

@@

@@

The pupillometry results are summarised in Fig. \@ref(fig:pupildata). The group average waveform for binocular presentation is shown in Fig. \@ref(fig:pupildata)a. There is a substantial pupil constriction at stimulus onset, followed by visible oscillations at the flicker frequency (2Hz, see waveform at foot). The average Fourier spectrum is displayed in Fig. \@ref(fig:pupildata)b, and shows a clear spike at 2 Hz, but no evidence of a second harmonic response at 4Hz. These results demonstrate that our paradigm can evoke measurable steady-state pupil responses at 2Hz.

@@

Fig. \@ref(fig:pupildata)c shows contrast response functions in response to stimuli flickering only at 2Hz. Response amplitudes increased monotonically with target contrast, confirming that our paradigm is suitable for measuring contrast-dependent differences in response (to our knowledge this is the first time this has been demonstrated). The amplitude of the binocular condition (blue squares) is consistently greater than that of the monocular condition (red circles) across all target contrasts. A $2\times5$ repeated measures $ANOVA^2_{circ}$ [@Baker2021] comparing these conditions revealed a significant main effect of target contrast (F(`r monbinPstats$dfA`,`r monbinPstats$dfR`) = `r round(monbinPstats$FratioA,digits=2)`, $p$ < 0.001), a significant effect of condition (F(`r monbinPstats$dfB`,`r monbinPstats$dfR`) = `r round(monbinPstats$FratioB,digits=2)`, $p$ < 0.001), and a significant interaction (F(`r monbinPstats$dfAB`,`r monbinPstats$dfR`) = `r round(monbinPstats$FratioAB,digits=2)`, $p$ < 0.001). The dichoptic condition begins at a much higher amplitude, owing to binocular combination of the target and high (48%) contrast mask, and then increases slightly with increasing target contrast (main effect of target contrast: F(`r dichA2c$dfM`,`r dichA2c$dfR`) = `r round(dichA2c$Fratio,digits=2)`, $p$ < `r round(dichA2c$pval,digits=3)`).

In Fig. \@ref(fig:pupildata)d, we plot responses to monocular target stimuli flickering at 2Hz, when the other eye viewed stimuli flickering at 1.6Hz (the red monocular-only data are replotted from Fig. \@ref(fig:pupildata)c for comparison). When the 1.6Hz component had the same contrast as the target (the binocular cross condition, shown in purple) responses were facilitated slightly at low contrasts, and suppressed at the highest target contrasts (interaction between contrast and condition: F(`r monbinPstatsX$dfAB`,`r monbinPstatsX$dfR`) = `r round(monbinPstatsX$FratioAB,digits=2)`, $p$ < 0.001). When the 1.6Hz component had a fixed contrast of 48% (the dichoptic cross condition, shown in yellow), responses were suppressed slightly across the contrast range (interaction between contrast and condition: F(`r mondichPstatsX$dfAB`,`r mondichPstatsX$dfR`) = `r round(mondichPstatsX$FratioAB,digits=2)`, $p$ < 0.001).

Fig. \@ref(fig:pupildata)e shows responses at 1.6Hz, for the same conditions, as well as for a condition in which a monocular stimulus flickered at 1.6Hz (grey circles). Surprisingly there again appears to be a slight facilitation effect in the binocular cross condition, particularly at lower contrasts. The dichoptic cross condition does not show clear modulation with target contrast.

Fig. \@ref(fig:EEGdata) shows equivalent results, measured contemporaneously using EEG. Fig. \@ref(fig:EEGdata)a shows the group average waveform for binocular presentation, and Fig. \@ref(fig:EEGdata)b shows the Fourier spectrum for binocular presentation, both averaged across four posterior electrodes (_Oz_, _POz_, _O1_ and _O2_, marked on the inset scalp plots). Unlike for the pupillometry data, there are clear responses at both the first harmonic frequency (2Hz), and also the second harmonic frequency (4Hz). We therefore calculated contrast response functions at both first and second harmonic frequencies.

@@

When stimuli in both eyes flicker at 2Hz, the binocular responses at the first (Fig. \@ref(fig:EEGdata)c) and second (Fig. \@ref(fig:EEGdata)f) harmonics are substantially greater than the monocular responses, particularly at high contrasts. Analysis of variance on the complex values ($ANOVA^2_{circ}$) revealed a main effect of contrast (F(`r monbinEstats$dfA`,`r monbinEstats$dfR`) = `r round(monbinEstats$FratioA,digits=2)`, $p$ < 0.001) and an interaction effect (F(`r monbinEstats$dfAB`,`r monbinEstats$dfR`) = `r round(monbinEstats$FratioAB,digits=2)`, $p$ < 0.001), but no effect of condition ($p$ = `r round(monbinEstats$pvalB,digits=2)`) at the first harmonic, with a similar pattern of results obtained at the second harmonic. For the cross-frequency conditions (Fig. \@ref(fig:EEGdata)d,g), there was no appreciable effect of adding a 1.6Hz component on the response at 2Hz or 4Hz (no effect of condition, and no interaction). Similarly, there were no clear interocular interactions between frequencies in the responses at 1.6Hz (Fig. \@ref(fig:EEGdata)e) and 3.2Hz (Fig. \@ref(fig:EEGdata)h). This pattern of results suggests that processing of temporal luminance modulations happens in a more linear way in visual cortex (indexed by EEG), compared with subcortical pathways (indexed by pupillometry), and shows no evidence of interocular suppression.

@@

Finally, we calculated the ratio of binocular to monocular responses across the three data types from Experiment 1. Fig. \@ref(fig:BSratios) shows that these ratios are approximately $\sqrt2$ across the low-to-intermediate contrast range for all three data types. At higher contrasts, we see ratios of 2 or higher for the EEG data, but much weaker ratios near 1 for the pupillometry data. Note that the ratios here are calculated on a per-participant basis and then averaged, rather than being the ratios of the average values shown in Fig. \@ref(fig:pupildata) and \@ref(fig:EEGdata). A $3 \times 5$ repeated measures ANOVA on the logarithmic (dB) ratios found a main effect of contrast (F(`r round(ratioaov$ANOVA$DFn[2]*gg[2],digits=2)`,`r round(ratioaov$ANOVA$DFd[2]*gg[2],digits=2)`) = `r round(ratioaov$ANOVA$F[2],digits=2)`, \emph{p} < `r round(ratioaov$ANOVA$p[2],digits=3)`), no effect of data modality (F(`r ratioaov$ANOVA$DFn[1]`,`r ratioaov$ANOVA$DFd[1]`) = `r round(ratioaov$ANOVA$F[1],digits=2)`, \emph{p} = `r round(ratioaov$ANOVA$p[1],digits=2)`), but a highly significant interaction (F(`r round(ratioaov$ANOVA$DFn[3]*gg[3],digits=2)`,`r round(ratioaov$ANOVA$DFd[3]*gg[3],digits=2)`) = `r round(ratioaov$ANOVA$F[3],digits=2)`, \emph{p} < `r round(ratioaov$ANOVA$p[3],digits=4)`).

@@

Results from the temporal frequency experiment are shown in Fig. \@ref(fig:TFdata). Fig. \@ref(fig:TFdata)a shows the Fourier spectra for responses to binocular flicker at 5 different frequencies (2, 4, 8, 16, and 30 Hz). From 2 to 16 Hz, clear signals are observed at each fundamental frequency, and typically also their higher harmonics (integer multiples of the fundamental). However, at 30 Hz (upper row), the responses recorded were not demonstrably above the noise baseline. Fig. \@ref(fig:TFdata)b compares the monocular and binocular responses at each stimulation frequency. Here we replicate the substantial summation effect across frequencies up to and including 16Hz (Fig. \@ref(fig:TFdata)c), demonstrating that strong binocular facilitation in the EEG data of Experiment 1 cannot be attributed to our use of 2Hz flicker.

@@

The results from the matching experiment are shown in Fig. \@ref(fig:matchingdata). Each data point indicates the contrast levels required in each eye that were perceptually equivalent to the binocular 24\% (red circles) and 48\% (blue circles) matching contrasts. At both matching contrasts, we see a very substantial increase in the physical contrast required for a monocular target (data points along the x- and y-axes), compared to a binocular target (points along the diagonal of x=y). For example with a 48\% match, the monocular targets required contrasts close to 100\%, whereas binocular targets required a contrast of around 50\%. The data points between these extremes also fall close to the predictions of a linear summation model (diagonal dotted lines), and are inconsistent with a winner-takes-all (or MAX) model (dashed lines). Overall, these matching results are consistent with the approximately linear summation effects observed in the EEG data of Experiment 1 (Fig. \@ref(fig:EEGdata)c,f).

@@

We fitted a computational model to the data from Experiments 1 & 3 using a hierarchical Bayesian approach. The model behaviour is displayed in Fig. \@ref(fig:modelfigure)e-h, with empirical data replotted in Fig. \@ref(fig:modelfigure)a-d for comparison. In general, the model captures the key characteristics of the empirical data. Note that there are some minor discrepancies, which are a consequence of the hierarchical nature of the modelling. In brief, the model is fitted to the amplitudes for each participant, and group-level parameter estimates are derived based on these fits (see Table \@ref(tab:paramtable)). This procedure discards the phase information, whereas the empirical averages are coherently averaged across participants (retaining phase information). This explains the amplitude differences between model and data, particularly at low target contrast levels, but is of little consequence for the pattern of relative responses across conditions, which is our main focus here.

We were particularly interested in comparing the weight of interocular suppression across data sets. We therefore plot the posterior distributions for this parameter for all four data sets (see Fig. \@ref(fig:modelfigure)i). The key finding is that the pupillometry results (green distribution) display a much greater weight of interocular suppression compared with the other data sets (grey, purple and yellow distributions). There is no overlap between the pupillometry distribution and any of the other three. All four distributions are also meaningfully below a weight of 1 -- the value that previous work using grating stimuli would predict [@Meese2006; @Baker2017], and the peak location of our prior distribution (black curve). These results offer an explanation of the empirical data: the strong interocular suppression for the pupillometry data is consistent with the weak binocular facilitation, and measurable dichoptic masking observed using that method. The weaker suppression for the other experiments is consistent with the near-linear binocular facilitation effects, and absent dichoptic masking.

@@

Classic studies investigating the neurophysiological architecture of V1 reported that cells in cytochrome-oxidase 'blobs' [@Horton1981; @Livingstone1984] are biased towards low spatial frequencies [@Tootell1988; @Edwards1995], and relatively insensitive to stimulus orientation [@Horton1981; @Livingstone1984; though see @Economides2011]. As the blob regions are embedded within ocular dominance columns [@Horton1981], they are also largely monocular [@Livingstone1984; @Tychsen2004]. More recent work has reported psychophysical evidence for unoriented chromatic [@Gheiratmand2013] and achromatic [@Meese2011] mechanisms, that also appear to be monocular. Our use of luminance flicker might preferentially stimulate these mechanisms, perhaps explaining why our EEG data show little evidence of binocular interactions. Indeed, our EEG results could potentially be explained by a model involving entirely non-interacting monocular channels, with the binocular facilitation effects we find (e.g. Fig. \@ref(fig:BSratios) & \@ref(fig:TFdata)) owing to additivity of the electrophysiological response across independent monocular cells, rather than requiring binocular neurons. However our matching data (Fig. \@ref(fig:matchingdata)), as well as everyday perceptual experience, indicates that luminance signals must be combined physiologically at some stage, and that this process still involves weak or absent interocular suppression.

@@

Thirty (20 females), twelve (7 females) and ten (3 females) adult participants, whose ages ranged from 18 to 45, were recruited for Experiments 1, 2 and 3 respectively. All participants had normal or corrected to normal binocular vision, and gave written informed consent in accordance to the Declaration of Helsinki. Our procedures were approved by the Ethics Committee of the Department of Psychology at the University of York (identification number 792).

The stimuli were two discs of achromatic flickering light with a diameter of 3.74 degrees, presented on a black background. The same stimuli were used for all three experiments. Four dark red lines were added around both discs to help with their fusion into one binocular disc (see insert in Fig. \@ref(fig:pupildata)b for an example of the fused stimulus). The discs were viewed through a four-mirror stereoscope, which used front silvered mirrors to avoid internal reflections, and meant that participants saw a single fused disc. The use of a stereoscope allowed us to modulate the stimuli in three different ocular configurations: monocular, binocular, and dichoptic. Note that during monocular presentation of flicker, the unstimulated eye still saw the static (non-flickering) disc of mean luminance.

We implemented the model within a Bayesian framework using the Stan software [@Carpenter2017]. This allowed us to estimate group-level posterior parameter distributions for the weight of interocular suppression, $w$, and the other free model parameters $R_{max}$, $Z$ and $n$. The prior distributions for all parameters were Gaussian, with means and standard deviations of 1 and 0.5 for $w$ and $R_{max}$, and 5 and 2 for $Z$ and $n$, with these values chosen based on previous literature [@Meese2006; @Baker2012]. We sampled from a Student's t-distribution for the amplitudes in the pupillometry and EEG experiments, and from a Bernoulli distribution for the single trial matching data. The models were fit using the individual data across all participants, independently for each data set. Because we used coherent averaging across participants, the group average amplitudes are shifted vertically relative to the model predictions, which are based on hierarchical fits to the individual participant amplitudes (put another way, the model does not implement coherent averaging). However it is clear that the character of the model in all cases gives a good representation of the data, as can be seen in Fig. \@ref(fig:modelfigure). We took posterior samples at over a million steps for each data set, using a computer cluster, and retained 10\% of samples for plotting.

## Preregistration

We initially preregistered our main hypotheses and analysis intentions for the first experiment. We then conducted a pilot study with N=12 participants, before making some minor changes to the stimulus (we added dim red lines to aid binocular fusion). We then ran the main experiment, followed by two additional experiments that were not preregistered. The preregistration document is available on the project repository: https://doi.org/10.17605/OSF.IO/TBEMA.

# Data availability

Raw data files are available on the project repository: https://doi.org/10.17605/OSF.IO/TBEMA.

# Code availability

Experimental and analysis codes are available on the project repository: https://doi.org/10.17605/OSF.IO/TBEMA.

# Funding

Supported by Biotechnology and Biological Sciences Research Council grant BB/V007580/1 awarded to DHB and ARW, and Wellcome Trust grant 213616/Z/18/Z to AB.

# Author information

## Authors and Affiliations

**Department of Psychology, University of York, York, UK**

Federico G. Segala, Aurelio Bruno, Myat T. Aung, Alex R. Wade &  Daniel H. Baker

## Contributions

**F.G.S.**: Methodology, software, formal analysis, investigation, data curation, writing - original draft, writing - review & editing, visualization. **A.B.**: Conceptualization, writing - review & editing, supervision, project administration, funding acquisition. **M.A.**: Software, resources, writing - review & editing. **A.R.W.**: Conceptualization, methodology, resources, writing - review & editing, supervision, project administration, funding acquisition. **D.H.B.**: Conceptualization, methodology, software, formal analysis, investigation, resources, data curation, writing - original draft, writing - review & editing, visualization, supervision, project administration, funding acquisition.

## Corresponding Author

Correspondance to Federico G. Segala.

# Ethics declarations

## Competing interests
